---
title: "Day Care Fines"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


```{r, warning=FALSE, message=FALSE}
library(mosaic)
library(DT)
library(pander)
library(car)
library(tidyverse)

# To get the "late" dataset into your R Console, 
# you need to go to your file menu of R-Studio and select
# "Session, Set working directory, To source file location"
# Then click the green "play" arrow in 
# the top right corner of this gray R-Chunk. 

## Wide data from original article:
late <- read.csv("../../Data/late.csv", header=TRUE) 

## Change to "long" data for use in R:
Late <- pivot_longer(late, #Start with wide data set late
               cols = starts_with("Week"), #Find columns to gather from long data
               names_to = "Week", #New column name of Weeks in long data
               names_prefix = "Week", #Remove prefix of Week from Week1, Week2, ...
               names_transform = list(Week= ~as.numeric(.x)), #Make Week a numeric column of data
               values_to = "NumberofLateChildren") #Name of data column in long data

```





## Background {.tabset .tabset-fade}

### Overview
<!-- Be sure to add your own background here before beginning the Analysis. -->



The overall question is: what is the overall effect of a fine for picking up your children late from daycare?

<!--End your background before here.-->
<!--Click the little black arrow to 
the left to expand/hide these next sections.-->
### Study Details

<div style="color:#a8a8a8;">
This background is quoted directly from the article ["A Fine is a Price"](http://rady.ucsd.edu/faculty/directory/gneezy/pub/docs/fine.pdf). 
</div>

<div style="padding-left:30px; padding-right:30px;">

There are two types of day-care centers in Israel: private and public. A study was conducted in 10 private day-care centers in the city of Haifa from January to June 1998. All of these centers are located in the same part of town, and there is no important difference among them. During the day children are organized into groups according to age, from 1 to 4 years old. Each day-care center is allowed to hold a maximum of 35 children. In some exceptional cases a few additional children are allowed. The fee for each child is NIS 1,400 per month. (The NIS is the New Israeli Shekel.) At the time of the study, a U.S. dollar was worth approximately NIS 3.68, so the fee was about $380 at that time. 

The contract signed at the beginning of the year states that the day-care
center operates between 0730 and 1600. There is no mention of what happens if parents come late to pick up their children. In particular, before the beginning of the study, there was no fine for coming late. When parents did not come on time, one of the teachers had to wait with the children concerned. Teachers would rotate in this task, which is considered part of the job of a teacher, a fact that is clearly explained when a teacher is hired. Parents rarely came after 1630. 

A natural option [to fix the problem of parents showing up late] is to introduce a fine: every time a parent comes late, [they] will have to pay a fine. Will that reduce the number of parents who come late? If the fine is removed, will things revert back to the way they were originally?

The overall period of the study was 20 weeks. In the first 4 weeks we
simply recorded the number of parents who arrived late each week. At the
beginning of the fifth week, we introduced a fine in six of the 10 day-care centers, which had been selected randomly. The announcement of the fine was made with a note posted on the bulletin board of the day-care center. Parents tend to look at this board every day, since important announcements are posted there. The announcement specified that the fine would be NIS 10 for a delay of 10 minutes or more. The fine was per child; thus, if parents had two children in the center and they came late, they had to pay NIS 20. Payment was made to the principal of the day-care center at the end of the month. Since monthly payments are made to the owner during the year, the fines were added to those amounts. The money was paid to the owner, rather then to the teacher who was staying late (and did not get any additional money). The teachers were informed of the fine but not of the study. Registering the names of parents who came late was a common practice in any case.

At the beginning of the seventeenth week, the fine was removed with no explanation. Notice of the cancellation was posted on the board. If parents asked why the fines were removed, the principals were instructed to reply that the fine had been a trial for a limited time and that the results of this trial were now being evaluated.

A comparison with other fines in Israel may give an idea of the size of
the penalty that was introduced. A fine of NIS 10 is relatively small but not insignificant. In comparison, the fine for illegal parking is NIS 75; the fine for driving through a red light is NIS 1,000 plus penalties; the fine for not collecting the droppings of a dog is NIS 360. For many of these violations, however, detection and enforcement are low or, as in the case of dog dirt, nonexistent in practice. A baby-sitter earns between NIS 15 and NIS 20 per hour. The average gross salary per month in Israel at the time of the study was NIS 5,595.

</div>


### The Data (Wide)


The `late` Day Care Center data is shown here in the "wide data format". 

```{r, out.width=300}
#Show the full width of the "Wide" version of the late data:
pander(late, split.tables = Inf)
```


### The Data (Long)

The `Late` Day Care Center data is shown here in the "long data format".

```{r}
pander(Late)
```


 
##  
<!-- You can begin writing again after here.-->



## Analysis

We will separate the data out into three different time periods, pre-fine, fine and post-fine. Pre-fine is the first four weeks of the study,fine is the twelve weeks following the pre-fine, and post-fine is the last four weeks of the study. We separate out the study to see if a fine really helps the situation, or does nothing.

### Hypothesis
Hypothesis 1:

Does adding a fine help the daycare's with the late parents problem?
$$
H_0: \mu_\text{pre-fine}=\mu_\text{fine}=\mu_\text{post-fine}
$$
$$
H_a: \mu_i \neq \mu \text{ for at least one } i \in\{1 = \text{pre-fine}, 2 = \text{fine}, 3 = \text{post-fine}\}
$$
Hypothesis 2:

Is there a difference between a daycare that has a fine and one that does not?
$$
H_0: \mu_\text{fine daycare} = \mu_\text{control daycare}
$$

$$
H_a: \mu_\text{fine daycare} \neq \mu_\text{control daycare}
$$
Hypothesis 3:

Is there an interaction between a daycare that implements a fine and a daycare that does not.
$$
H_0: \text{The effect of a fine is the same for all daycares.}
$$
$$
H_a: \text{The effect of a fine is not the same for all daycares.}
$$


A significance level of $\alpha = 0.05$ will be used for this study.

### Two-Way ANOVA


For this analysis we will be performing the following Two-way ANOVA.

```{r}
Late$FinePeriod <- cut(Late$Week, c(0,4,17,20), labels = c("Pre-Fine","Fine","Post-Fine"), ordered_result = TRUE)

fine.aov <- aov(NumberofLateChildren ~ FinePeriod + Treatment+FinePeriod:Treatment, data=Late)
summary(fine.aov) %>% pander()
```


The ANOVA table above has three different p-values, one for each individual hypothesis listed above. We can conclude that there is sufficient evidence to reject the null hypothesis on all of the three test that were performed. For the first test on if adding a fine will help with the late parents problem, we see that $p=2.618e-06$. For the second test on if there was a difference between a daycare that has a fine for late parents and one that does not,  we see that $p=4.275e-12$. For the third test on if there is an interaction between a daycare that implements a fine and a daycare that does not, we see that $p=4.263e-06$

### Diagnostic Plots

The ANOVA test is somewhat questionable as shown in the plots below. From the Normal Q-Q plot we can assume that it is somewhat normal, but with the constant variance assumption (Residuals vs Fitted) is questionable. This is because it appears to get larger as you continue on to the left; however, the change in variance is not substantial enough to reject the ANOVA test, so we can assume the results as valid.
```{r}
par(mfrow=c(1,2))
plot(fine.aov, which=1:2, pch=16)
```




### Graphical Summaries

Below are graphics to help understand each individual test that ran.

#### Fine Periods

This graph shows that there is an increase in number of late children at the daycares from Pre-Fine period to Fine period and from Fine period to Post-Fine period. We can see that the biggest increase is between the Pre-Fine period which had an average of 8.8 late children where the Fine period had an average of 13.47 late children. Then we a small increase from the Fine period to the Post-Fine period with an average of 15.23 late children. With the help from the graph we can understand why the p-value (2.618e-06) indicates that there is sufficient evidence to reject the null hypothesis.
```{r}
xyplot(NumberofLateChildren ~ FinePeriod, data=Late, type=c("p","a"), main="Significance of a Fine", xlab = "Fine Period",ylab = "Number of Late Children",col='cyan4') 

Late %>%
  group_by(FinePeriod) %>%
  summarise(`Mean Number of Late Children`=mean(NumberofLateChildren)) %>%
  pander(caption="Mean Number of Late Children according to the Fine Period")

```




#### Daycares that offer Fines vs Daycares that do not offer Fines

From this graph we can see that there is an increase between the number of late children and the daycares that implement a fine and the daycares that do not implement a fine. The average number of late children for the daycares that do not implement a fine is 9.188. The average number of late children for the daycares that implement a fine is 15.21. With the help from the graph we can understand why the p-value (4.275e-12) indicates that there is sufficient evidence to reject the null hypothesis.

```{r}
xyplot(NumberofLateChildren ~ as.factor(Treatment), data=Late, type=c("p","a"), main="Significance of Treatment",xlab = "Treatment",ylab = "Number of Late Children", col='aquamarine3') 

Late %>%
  group_by(Treatment) %>%
  summarise(`Mean Number of Late Children`=mean(NumberofLateChildren)) %>%
  pander(caption="Mean Number of Late Children according to the Type of daycare")

```


#### Type of Daycare Depending on Fine Period

This graph puts the two previous graphs together, which helps us understand why the  p-value (4.263e-06) indicates that there is sufficient evidence to reject the null hypothesis. From this graph we can see that there are more late children on average in the daycares that do no implement a fine (10) than the daycares that do implement a fine (8), during the Pre-fine period. After the Pre-Fine period, we see that there are more late children in the daycares that do offer a fine than the daycares that do not implement a fine.


```{r,warning=FALSE, message=FALSE}
xyplot(NumberofLateChildren ~ FinePeriod, data=Late, groups=Treatment, type=c("p","a"), main="Significance of the Interaction",xlab = "Fine Period",ylab = "Number of Late Children", auto.key=list(corner=c(1,1)))

Late %>%
  group_by(FinePeriod, Treatment) %>%
  summarise(ave=mean(NumberofLateChildren),.groups="keep") %>%
  spread(FinePeriod, ave) %>%
  pander(caption="Mean Number of Late Children according to Treatment (Control,Fine) and Fine Period (Pre-Fine, Fine, Post-Fine)")
```




### Conclusion

In general, we recommend using a daycare that does not implement a fine rather than using a daycare that does implement a fine. One reason for this conclusion is because you will not be receiving a fine if you have to pick up your child late. Another reason why we recommend using a daycare that does not implement a fine is because that there are significantly lower number of children being picked up late (9.188) than the daycares that do implement a fine (15.21).
