---
title: "Chi Squared Test"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
---


<!-- Directions

Come up with a chi squared test analysis of your own creation. 

* Write the Background and clearly state your question. Then collect data and perform the chi-squared test analysis in order to answer your question. Use the proper order of first, ask the question, then second, figure out the answer.

* You can use data from 
    - An R data set
    - the internet (like weather data, Facebook data, or sports data) 
    - your own experiment 
    - your own observational study
* However,     
    - Do not violate the "Human Subjects Rules" of the university: http://www.byui.edu/executive-strategy-and-planning/institutional-research/research-and-assessments/research-and-surveys

* You need a sample size large enough to meet the requirements of the chi squared test. Unfortunately, the sample size is completely dependent on the situation you select. A good rule of thumb is 10 observations per "combination-group" in your data.

-->


<!--
"ownguns" & "region"
-->

```{r, message = FALSE, warning= FALSE}
library(readr)
library(tidyverse)
library(mosaic)
library(pander)
GSS2012 <- read_delim("../../Data/GSS2012.csv", 
                      "\t", escape_double = FALSE, trim_ws = TRUE)
```

## Background
A general survey was taken in 2012. Two questions from the survey caught my eye. "Do you happen to have in your home (or garage) any guns or revolvers?" and "Region of the Interview". We will look at the responses of having a gun in your home Yes (1) or No (2) and their region.

## Hypothesis
The question for our data is: Does the region of the country contribute having a gun in the home? In other words, is having a gun in your home and region of the country associated?

$$
H_0: \text{Gun in the home and Region of the country are independent}
$$
$$
H_a: \text{Gun in the home and Region of the country are associated (not independent)}
$$
$$
\text{Significance level: } \alpha= 0.05
$$

## Analysis

Here is a table that counts all the responses people entered in according to their responses to if they have a gun in their house and what part of the country they took the survey.
```{r}
GS <- select(GSS2012, c(owngun, region))
GS<- filter(GS, owngun %in% c(1,2))
GS1 <-GS %>%
  mutate(Yes_or_No =
   case_when(
      GS$owngun %in% 1 ~ "Yes",
      GS$owngun %in% 2 ~ "No"),
   Region = case_when(
     GS$region %in% 1 ~ "New England",
     GS$region %in% 2 ~ "Middle Atlantic",
     GS$region %in% 3 ~ "E. Nor. Central",
     GS$region %in% 4 ~ "W. Nor. Central",
     GS$region %in% 5 ~ "South Atlantic",
     GS$region %in% 6 ~ "E. Sou. Central",
     GS$region %in% 7 ~ "W. Sou. Central",
     GS$region %in% 8 ~ "Mountain",
     GS$region %in% 9 ~ "Pacific",
   ))
survtable<-table(GS1$Yes_or_No,GS1$Region)
pander(survtable)
```

From this table we can see that there are two groups that have answered No to owning a gun in their home and they are located in the E. Nor. Central and Pacific regions. The largest number of people that answered Yes (103) are located in the South Atlantic region.

```{r}
barplot(survtable, beside=TRUE, col=c("firebrick1","dodgerblue"),legend.text=TRUE,args.legend = list(x="topleft",bty="n",title="Gun in the Home"),xlab="Region",main="The Number of People that Keep Guns In Their Homes by Region \n Accourding to a 2012 Survey")
```

This Bar plot gives a better representation of what the data table shows. However this doesn't answer our question if these two variables are independent or associated. To answer this we will run a Chi-Squared Test. We will also look at the expected counts to see if the assumptions for this test are met.

```{r}
GSChi<-chisq.test(survtable)
pander(GSChi)
```
```{r}
pander(GSChi$expected)
```

All of the expected counts are larger than 10, so the requirements are met. 

Our results show a $P$-value of 2.514e-07. This $P$- value is smaller than out Significance level ($\alpha = 0.05$) so we must reject the null hypothesis.

Here are  our residuals. This will let us see how much our observed counts differ from the expected counts if our null hypothesis were true.
```{r}
pander(GSChi$residuals)
```

The greatest difference between observed and expected counts was 3.346. This probably contributed most to our resulting p-value.

## Conclusion

Since we have to reject our null hypothesis we can conclude that the number of people that keep a gun in their home and their region are associated and not independent. This means we can assume that certain regions of the United States has an affect if people have a gun in their home.


