---
title: "Rent"
output: 
  html_document:
    theme: cerulean
    code_folding: hide
    toc: true
    toc_float: true
---

<!-- Welcome to the first Analysis for Math 325! 
The goal of this first assignment is to have you produce a graphical summary and a numerical summary of the "Rent" data set. The fun part is that you get to decide what type of graph and what type of numerical summary to create! You will need to read through the "green text" instructions of this document to find the places where you are supposed to enter R-code in order to complete the assignment. -->



```{r, message=FALSE, warning=FALSE}
# This gray area is called an "R-chunk". 
# These library commands install some powerful functions for your use later on.
library(mosaic)
library(pander)
library(tidyverse)
library(DT) 
library(plotly)

# This read_csv command reads in the "Rent" data set into an object called "Rent"
Rent <- read_csv("../Data/Rent.csv")

# To load this data set into your R-Console do the following:
#  1. From your top file menu select "Session -> Set Working Directory -> To Source File Location"
#  2. Press the green "play button" in the top right corner of this gray box (which is called an "R-chunk").
#  3. Then in your "Console" window of 
```

<br />

## Background

Here is a data table showing the available approved housing apartment options at BYU-Idaho for single students. There are 122 entries comprising 57 female and 65 male apartment options. Is it cheaper for a Private room or a Normal room, and is it more affordable than other universities?
I believe that Normal rooms will be cheaper than the private rooms, but both will be cheaper than the rent at any other university.

```{r, warning=FALSE}
# Code to get you started. 
# View(...) works great in the Console, but datatable(...) must be
# used instead within an R-chunk.
datatable(Rent, options=list(lengthMenu = c(3,10,30)), extensions="Responsive")
```


## Graphic

<!-- Present an interesting graphic using the Rent data set. Do not create the same graphics that are shown in the [Good Example Analysis](./Analyses/StudentHousing.html), but you can certainly use those graphics as a guide to inspire you on what to do yourself. -->

The scatterplot below is a representation of the relationship between private room rent and normal rent for a single semester at Brigham Young University-Idaho.


```{r, message=FALSE, warning=FALSE}
# Use this R-chunk to...
# ...Create an interesting graphic using the Rent data set.

plot_ly(Rent, x= Rent$PrivateRoomPrice, y= Rent$Price, size= Rent$FloorPlans, text= paste("Apartment Name:", Rent$Apartment,"\n", "Phone Number:", Rent$Phone, "\n", "Website:", Rent$Website), color= Rent$Gender, colors=c("deeppink","darkslategray3")) %>% layout(title=("Private Room Rent vs. Normal Rent at BYU-I"), xaxis=list(title="Private Room Rent (in dollars)"), yaxis=list(title="Normal Rent (in dollars)" ))

```


<!-- Write a few statements here that explain your graphic and what it shows. -->

From this scatterplot we can see a positive relationship between private room rent and normal rent. As private room rent increases so does the normal rent. We can also see from the graph that there are more apartments for women between 1400-1600 dollars in private room rent than there are for men. Another observation we can make is that there are more apartments for men between 800-1000 dollars in private room rent than women. 




```{r}
# Use this R-chunk to...
# ...compute and display a meaningful table of numerical summaries supporting your above graphic.
 outputTable <- rbind('Private Room Rent'= summary(Rent$PrivateRoomPrice), 'Normal Rent' = summary(Rent$Price))
outputTable2 <- rbind('r' = cor(Rent$PrivateRoomPrice,Rent$Price,use= "complete.obs"))
pander(outputTable,caption = "Private Room Rent vs. Normal Rent at Brigham Young University-Idaho")
pander(outputTable2)
```

<!-- Write a few statements here that introduce your table of numerical summaries and what they show. -->

The data table above compares the quartiles of both private room rent and normal rent. From this data we can come to the conclusion that there is not a big difference between private room rent and normal rent in the first quartile, but as we get to quartile three, we can see that there is a significant difference between the two prices. Then the r shows the correlation between private room rent and normal rent. Since the r value is close to one, then we can come to the conclusion that there is a positive correlation between private room rent and normal rent.


## Conclusion

According to debt.org, for a normal college student that decides to live in an apartment will have to pay $675 per month. One thing about Brigham Young University - Idaho is that it is a three semester college, meaning that there are three semesters in one school year. In a normal two semester college, there are about 6 months in a semester, so the cost of a college student to rent an apartment would be about $4,050 per semester. The mean for the price for a private room per semester at Brigham Young University-Idaho is $1,165. From this data, we can come to the conclusion that private room's rent is going to be more than what the normal room's rent is, and that Brigham Young University-Idaho is more affordable for housing per semester than what the average says about other colleges.




